site-wise niche-overlap by elevation

### 2020-11-26 analyses/old_scripts/niche_relations_v3.Rmd
did so many things

Load packages 
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(ggbipart)
library(GGally)
library(vegan)
library(ggvegan)
library(ggrepel)
library(gridExtra)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, 
                   elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, 
                             levels = c("unten", "mitte", "oben")),
         elev.class2 = factor(elev.class2, 
                             levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  mutate(pbl.fac = factor(if_else(bb.sp %in% c("hort", "pasc", "wurf"), "long", "short")))

# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  #left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>%

  # Fill in missing k.type scores
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) %>%

  # Fill in missing k.type scores
  mutate(k.type.s = if_else(plant.genus == "Phyteuma", 7.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Phyteuma", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Aconitum", 5.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Aconitum", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Cardamine", 1.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Cardamine", 1, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Euphrasia", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Euphrasia", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Lamium", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Lamium", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Mentha", 2.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Mentha", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Rumex", 0.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Rumex", 0, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Salix", 9.0, k.type.s),
         k.type.ss = if_else(plant.genus == "Salix", 9, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Silene", 4.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Silene", 4, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Stachys", 5.1, k.type.s),
         k.type.ss = if_else(plant.genus == "Stachys", 5, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.genus == "Taraxacum", 7.2, k.type.s),
         k.type.ss = if_else(plant.genus == "Taraxacum", 7, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana aspera", 2.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana aspera", 2, k.type.ss)) %>%
  mutate(k.type.s = if_else(plant.sp == "Gentiana ciliata", 4.1, k.type.s),
         k.type.ss = if_else(plant.sp == "Gentiana ciliata", 4, k.type.ss))

### Auxiliary data, i.e. all the covariates we need to control for
BB_covars <- net %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp)),
            sqrt.bb.abund = sqrt(n()))

FL_covars <- survey %>%
  mutate(year = factor(year(date))) %>%
  filter(flower.cover > 0) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp)),
            log.fl.abund = log(sum(flower.cover)))

### Species-level by site (we will also separate pollen and nectar foragers)
BB_covars.splevel <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
  group_by(bb.sp, site, year, pollen) %>%
  summarize(bb.sp.abund = n())

### Species names and such for plotting
species_names <- cbind(c("bss", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit", 
                         "humi"),
                       c("B. sensu-strictu", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus", 
                         "B. humilis"),
                       c("B. ss.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit.", 
                         "B. humi"),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "minor", "common", 
                         "minor")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4) %>%
  mutate(focal = if_else(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE))
```

@@

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  #group_by(site, date, elev.mean, bb.sp, year) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  #group_by(site, elev.mean, bb.sp, year) %>%
  #summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("./output/bb_range.pdf", height = 6, width = 6)

bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  left_join(species_names)

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Functions
```{r, echo = FALSE}
# Add BB trait data for envfit() call inside map2()
trait_join <- function(x, y = bb_traits) {
  x %>%
    select(bb.sp) %>%
    left_join(y) %>%
    column_to_rownames(var = "bb.sp") %>%
    select(pbl.w)
}

# Fortify envfit() object fitted to BB tongue length data
envfit_fort <- function(x){
  fortify(x) %>%
  select(Score = Type, Label, MDS1, MDS2) %>%
  mutate(Label = case_when(Label == "pbl.cat" ~ "T.class"),
         Score = case_when(Score == "Vector" ~ "T.class"))
}

# Fortify PCOA 
pcoa_fort <- function(x){
  fortify(x, axes = 1:2) %>%
  mutate(Score = case_when(
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp"))
}

# Plot PCOA; to avoid overplotting, loadings < min_load are not plotted
pcoa_plot <- function(x, min_load = 0.25) {

  t.class <- x %>%
    filter(Score == "BB.sp") %>%
    mutate(T.class = factor(if_else(Label %in% c("hort", "pasc", "wurf"), "long", "short")))

  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), 
                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) +
    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
    theme_classic()
}

# Niche analysis function; generates PCOA, envfit, ordination plot, and tongue length R2 coming out of envfit
niche_analysis1 <- function(x, y, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen) %>%
    nest() %>%
    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(comm2, function(x) x %>% 
                          left_join(y) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac)),
           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

# This version break it down by elev.class
niche_analysis2 <- function(x, y, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(elev.class, bb.sp, !!criterion, pollen) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(elev.class, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class) %>%
    nest() %>%
    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(comm2, function(x) x %>% 
                          left_join(y) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac)),
           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, elev.class, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

# ...and this version breaks it down by elev.class and period
niche_analysis3 <- function(x, y, criterion) {

  criterion <- enquo(criterion)

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>%
    filter(!is.na(pollen)) %>%
    group_by(elev.class) %>%
    mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) %>%
    group_by(elev.class, period, pollen, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site
    group_by(elev.class, period, bb.sp, pollen) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair per site-date
    dplyr::select(-freq) %>%
    group_by(pollen, elev.class, period) %>%
    nest() %>%
    mutate(comm1 = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop)), # spread into the kind of array vegan likes
           comm2 = map(comm1, function(x) replace(x, is.na(x), 0)),
           comm3 = map(comm2, function(x) column_to_rownames(x, var = "bb.sp")),
           pcoa = map(comm3,  function(x) capscale(x ~ 1, distance = "horn")),
           traits = map(comm2, function(x) x %>% 
                          left_join(y) %>% 
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac)),
           adonis = map2(comm3, traits, function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           tongue.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]),
           tongue.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]]),
           MDS1.varexp = map(pcoa, function(x) x$CA$eig[[1]]),
           MDS2.varexp = map(pcoa, function(x) x$CA$eig[[2]])) %>%
    unnest(cols = c(tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp)) %>%
    select(pollen, elev.class, period, comm2, comm3, traits, pcoa, pcoa.fort, adonis,
           pcoa.plot, tongue.r2, tongue.p, MDS1.varexp, MDS2.varexp) 
}

@@

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair),
           pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
    left_join(BB_covars) %>% # add BB covars
    left_join(FL_covars) %>% # add floral covars
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}
```

@@

To avoid overplotting plant species / Kugler classes, I have only plotted species loading above a given threshold on either the first or second axis. Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet. A species that is visited heavily by all BB species will fall close to (0,0) and will therefore not be plotted.

@@

```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche.tax <- niche_analysis1(net, bb_traits, criterion = plant.sp.abb)

@@

# By ktype
niche.morph <- niche_analysis1(net, bb_traits, criterion = k.type.s)

@@

2.2 At mid elevation, a bunch of morphological traits come into play. As in the aggregate analysis, long-tonguers are associated with flag flowers (Kugler 6.1; *Trifolium pratense*, *Anthyllis vulneraria*) and the true lip flower (Kugler 5.1) *Rhinanthus glacialis*. Short-tongue species, however, are spread along a wide and almost linear gradient with *B. sensu-strictu* on one end, *B. soroensis* on the other, and *B. pratorum* in the middle. Toward the *bss* end of the gradient, we have the disk/bowl flower (Kugler 1.2) *Hypericum maculatum* (actually K 1.1, but I think 1.1 and 1.2 loaded very close to each other, but 1.1 was dropped by the overplotting threshold) and the unclassified flower (Kugler 0) *Plantago lanceolata*; I think unclassified flowers are those that Kugler reckoned to be wind-pollinated. On the *B. soroensis* end of the gradient, we have the flower head (Kugler 7.1) *Phyteuma orbiculare*. In the middle, near *B. pratorum*, we have the bell flower (Kugler 3.2) *Geum rivale*.

@@

```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche_x_elev.tax <- niche_analysis2(net, bb_traits, plant.sp.abb)

@@

# By ktype
niche_x_elev.morph <- niche_analysis2(net, bb_traits, k.type.s)

@@

# By ktype
niche_x_elevperiod.morph <- niche_analysis3(net, bb_traits, k.type.s)

@@

# Plot
## Binary
beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

>>>>

Load packages 
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
#library(GGally)
library(vegan)
library(ggvegan)
library(ggrepel)
#library(gridExtra)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("unten", "mitte", "oben"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  mutate(pbl.fac = factor(if_else(
    bb.sp %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
    )

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>% # add site data
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>% # add floral traits
  group_by(elev.class) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early", "mid", "late"))) # divide each elev.class into 3 equal time slices

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) # add floral traits
```

@@

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  mutate(focal = if_else(
    bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE)
    )

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("./output/bb_range_year.pdf", height = 6, width = 6)
```

Functions
```{r, echo = FALSE}
# Convert PCoA to tidy array ("fortify")
pcoa_fort <- function(x){
  fortify(x, axes = 1:2) %>% 
  mutate(Score = case_when( # rename default scores to match my data
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp"))
}

# Plot PCoA; to avoid overplotting, scores < min_score are not plotted
pcoa_plot <- function(x, min_score = 0.25) {

  # Add T.class variable for the BB.sp subset of the PCoA
  t.class <- x %>%
    filter(Score == "BB.sp") %>%
    mutate(T.class = factor(if_else(
      Label %in% c("hort", "pasc", "wurf", "gers", "mend", "hypn"), "long", "short"))
      )

  ggplot(filter(x, (abs(MDS1) > min_score | abs(MDS2) > min_score) | (Score == "BB.sp"))) + # no threshold for BB spp
    geom_polygon(data = t.class, aes(MDS1, MDS2, linetype = T.class), # draw polygons around T classes
                 color = "gray70", fill = NA, size = 1.25, key_glyph = "abline") +
    geom_point(aes(MDS1, MDS2, color = Score), size = 3, alpha = 0.75) + # add points
    ggrepel::geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") + # add labels
    theme_classic()
}

# Generates PCoAs, ADONIS tests, and ordination plots (fully aggregated)
## The ... represents the variables we will use to parse our analysis. For aggregate analysis, just leave the ...
## unpopulated, e.g. niche_analysis(net, bb_traits, plant.sp). 
## To parse by elev.class: niche_analysis(net, bb_traits, plant.sp)
## To parse by  elev.class, period, and pollen: niche_analysis(net, bb_traits, plant.sp, period, pollen)
## You get the idea. You should anyway -- you wrote it. https://tidyeval.tidyverse.org/multiple.html
niche_analysis <- function(x, y, criterion, ...) {

  criterion <- enquo(criterion) # this argument toggles taxonomic (plant.sp) vs. morphological (k.type.s) analysis

  x %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(..., bb.sp) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    group_by(...) %>%
    nest() %>% # nest into list-column
    mutate(comm1 = map(data, # spread into the kind of array vegan likes
                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
                         replace(is.na(.), 0)
                       ), 
           comm2 = map(comm1, # column to rowname
                       function(x) column_to_rownames(x, var = "bb.sp")
                       ),
           pcoa = map(comm2, # call PCoA
                      function(x) capscale(x ~ 1, distance = "horn")
                      ),
           traits = map(comm1, # join BB traits
                        function(x) x %>% 
                          left_join(y) %>% # y is the trait table, an argument in the main function call
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.fac) # pbl.fac is the variable I'm interested in
                        ),
           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
                         function(x, y) adonis(x ~ pbl.fac, data = y, method = "horn")
                         ),
           betadis = map2(comm2, traits, # betadisper to test for difference in dispersion between tongue classes
                         function(x, y) x %>%
                           vegdist(method = "horn") %>% # betadisper runs a PCoA under the hood; match method to my PCoA
                           betadisper(group = y$pbl.fac) 
                         ),
           betadis.permutest = map(betadis, permutest), # perumutation test
           betadis.p = map(betadis.permutest, # extract betadisper p-value
                           function(x) x$tab$'Pr(>F)'[[1]]
                           ),
           betadis.boxplot = map(betadis, boxplot), # boxplot of betadisper
           pcoa.fort = map(pcoa, pcoa_fort), 
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           adonis.r2 = map(adonis, # extract adonis r2
                           function(x) x$aov.tab$R2[[1]]
                           ),
           adonis.p = map(adonis, # extract adonis p-value
                          function(x) x$aov.tab$'Pr(>F)'[[1]]
                          ),
           MDS1 = map(pcoa, # extract eigenvalue for MDS1
                      function(x) x$CA$eig[[1]]
                      ), 
           MDS2 = map(pcoa, # extract eigenvalue for MDS2
                      function(x) x$CA$eig[[2]]
                      )) %>%
    unnest(cols = c(adonis.r2, adonis.p, betadis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
    select(..., pcoa, betadis, betadis.permutest, betadis.p, adonis, 
           adonis.r2, adonis.p, betadis.boxplot, pcoa.plot, MDS1, MDS2) 
}

@@

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair),
           pollen = factor(pollen)) %>% # ...and make it a factor so that mgcv can handle it
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}
```

@@

To avoid overplotting plant species / Kugler classes, I have only plotted species scoring above a given threshold on either the first or second axis. Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet. A species that is visited heavily by all BB species will fall close to (0,0) and will therefore not be plotted.

@@

```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb)

@@

# By ktype
niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.s)

@@

2.2 At mid elevation, a bunch of morphological traits come into play. As in the aggregate analysis, long-tonguers are associated with flag flowers (Kugler 6.1; *Trifolium pratense*, *Anthyllis vulneraria*) and the true lip flower (Kugler 5.1) *Rhinanthus glacialis*. Short-tongue species, however, are spread along a wide and almost linear gradient with *B. sensu-strictu* on one end, *B. soroensis* on the other, and *B. pratorum* in the middle. Toward the *bss* end of the gradient, we have the disk/bowl flower (Kugler 1.2) *Hypericum maculatum* (actually K 1.1, but I think 1.1 and 1.2 scored very close to each other, but 1.1 was dropped by the overplotting threshold) and the unclassified flower (Kugler 0) *Plantago lanceolata*; I think unclassified flowers are those that Kugler reckoned to be wind-pollinated. On the *B. soroensis* end of the gradient, we have the flower head (Kugler 7.1) *Phyteuma orbiculare*. In the middle, near *B. pratorum*, we have the bell flower (Kugler 3.2) *Geum rivale*.

@@

```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class)

@@

# By ktype
niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class)

@@

# By ktype
niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class, period)

@@

# Plot
## Binary
beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot.bin <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")
